BOLD Parameter Estimation Using Particle Filters

### 2010-08-16 doc/thesis/conclusion.tex
more on the results. Updated the HRF and applied pierces recommendations
"done"

\chapter{Conclusion}
\label{sec:Conclusion}
This work has demonstrated the use of the particle filter to
learn parameters of the BOLD model. Since the inception of the
BOLD equations, many attempts have been made use FMRI data to 
to learn these parameters. These attempts have typically either been extremely 
slow or relied on extensive assumptions. While the particle filter method
is not quick, 60 seconds to analyze each voxel is well within the capabilities
of the typical research lab. Previous attempts have also treated
the problem as if there were a single solution.

One significant finding of this work, that would not be clear without 
calculating a full posterior distribution, is the interplay 
between the parameters. The results of simulations clearly demonstrate
that identifying a single set of parameters is not possible with this 
model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
at this, the current work clearly demonstrates this fact. Therefore,
any single estimate of parameters is insufficient for analysis. As such,
approaches to the BOLD model that do not treat the parameters as distributions
will not be able to overcome the inherent indeterminability of the
parameters. Besides the Unscented Kalman Filter, this is the only approach
that accomplishes this task without repeatedly calculating the parameters 
to build a distribution. The Unscented Kalman Filter, of course, is limited
to Gaussian estimation which limits its ability to estimate the posterior.

The primary reason this method has not been used before is the high 
dimensionality of the system. In \cite{Murray2008} the idea of learning
the parameters is floated as the correct method, yet 
learning 7 parameters with a Monte Carlo method is often intractable. The 
concern was that so many parameters creates a prohibitively large 
search space, that requires too many particle to learn properly. To overcome
this difficulty, instead of starting the algorithm with 1000 particles,
the initial number of particles was set to 16,000. This meant that
when the search space was the largest, the number of particles was
sufficiently dense to represent the prior distribution. Then resampling
was performed for the first time, the number of particles was dropped, 
since the weight of most the particles had dropped to 0. The result is
a particle filter algorithm that spends computing resources only where
they are really needed.

Many of the approaches to determining areas of neural activation aim
to be Bayesian, yet ultimately must make limiting assumptions about
the distributions that turn the approach into classical statistics. 
This approach is a true Bayesian non-parametric approach that given
enough measurements will approach the true probability distribution
of the parameters. While the conclusion that the parameters are not
uniquely identifiable is a disappointing one in light years of research 
attempting to learn these parameters, it in fact further demonstrates
the need to estimate probability distribution rather than a single 
parameter estimate. At the same time, the output of the particle filter
is still about to give good estimates of the BOLD output, and not
dependent on heavy Gaussian smoothing. Therefore, all the current 
experimental paradigms to determine regional activity are still viable.
To borrow a computer term, the particle filter algorithm is backward 
compatible with the current methods.

>>>>

\chapter{Conclusions}
\label{sec:Conclusion}
This work has demonstrated the use of the particle filter to
learn parameters of the BOLD model. Since the inception of the
BOLD equations, many attempts have been made use FMRI data to 
to learn these parameters. These attempts have typically either been extremely 
slow or relied on extensive assumptions. While the particle filter method
is not quick, 40 seconds to analyze each voxel is well within the capabilities
of the typical research lab. Previous attempts have also treated
the problem as if there were a single solution.

One significant finding of this work, that would not be clear without 
calculating a full posterior distribution, is the interplay 
between the parameters. The results of simulations clearly demonstrate
that identifying a single set of parameters is not possible with this 
model. Although sensitivity tests in Deneux et al. certainly hinted
at this, the current work clearly demonstrates this fact \cite{Deneux2006}. Therefore,
any single estimate of parameters is insufficient for analysis. As such,
identification of the BOLD model that do not treat the parameters as distributions
will not be able to overcome the inherent indeterminability of the
parameters. Besides the Unscented Kalman Filter, this is the only approach
that accomplishes this task without repeatedly calculating the parameters 
to build a distribution. The Unscented Kalman Filter, of course, is limited
to Gaussian estimation which limits its ability to estimate the posterior.

The primary reason this method has not been used before is the high 
dimensionality of the system. In Murray, 2008 the idea of learning
the parameters is floated as the correct method, yet 
learning 7 parameters with a Monte Carlo method was deemed intractable
\cite{Murray2008}. The 
concern was that so many parameters creates a prohibitively large 
search space, that requires too many particle to learn properly. To overcome
this difficulty, instead of starting the algorithm with 1000 particles,
the initial number of particles was set to 16,000. This meant that
when the search space was the largest, the number of particles was
sufficiently dense to represent the prior distribution. Then resampling
was performed for the first time, the number of particles was dropped, 
since the weight of most the particles had dropped to 0. The result is
a particle filter algorithm that spends computing resources only where
they are really needed.

Many of the methods to determining areas of neural activation aim
to be Bayesian, yet ultimately must make limiting assumptions about
the distributions that remove many of the benefits. 
The particle filter is a true Bayesian non-parametric algorithm that, given
enough measurements will approach the true probability distribution
of the parameters. While the conclusion that the parameters are not
uniquely identifiable is a disappointing one in light years of research 
attempting to learn these them, it in fact further demonstrates
the need to estimate probability distribution rather than a single 
parameter estimate. At the same time, the output of the particle filter
is still able to give good estimates of the BOLD output, and is not
dependent on heavy Gaussian smoothing. Therefore, all the current 
experimental paradigms to determine regional activity are still viable.
To borrow a computer term, the particle filter algorithm is backward 
compatible with the current methods. Indeed, the particle filter excelled
at determining activation in very low SNR simulations 
(\autoref{sec:Multi-voxel Simulation}). Not only that, but the particle
filter algorithm identified areas of activation that were completely
missed by SPM (\autoref{fig:comp6}).

### 2010-08-16 doc/thesis/discussion.tex
"done"

\chapter{Discussion}
\label{sec:Discussion}
\section{Review of Results}
% Overview of results
%% Parameters under-constrained
%% Output estimates good
There are two possible, but unlikely causes for the large variance in
parameter estimates. First, it could be that the particle filter is
incapable of learning the model. However, given the high quality
of BOLD estimates, this is almost certainly not the case. The other 
possible explanation is that the given stimulus is not sufficiently
varied to bring out all the properties of the system. This also is unlikely
because of results in \autoref{sec:VeryLongSim} which used a random impulse
stimulus sequence. The high correlation of the parameters in \autoref{sec:VeryLongSim},
point to a significant trade off in parameters.

Thus the results show that the parameters of the BOLD model are under
constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
it is a notable conclusion \cite{Deneux2006}. Thus, other methods
that depend on a point estimate of the parameters, such as least squares 
or Kalman filters (which uses the first two moments) are limited. While estimates of
the  BOLD signal may still be correct, the estimate
of the underlying parameters and state variables cannot be described without using
a joint probability distribution function. In this sense, particle 
filters represent an important step forward in BOLD parameter 
estimation. Representing uncertainty with a mean
and variance is insufficient; so using a particle filter
or Bayesian estimate of the posterior is not simply an enhancement, 
but a necessary precaution.

Because of the ill-defined nature of the parameters, the results 
of Friston et al. may have been wider than true parameter distribution
\cite{Friston2002}. Any algorithm that attempts to calculate parameters
of the BOLD model should be run multiple times, to check for consistency.

As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information, closely resembled the results of SPM. While the activation
tests were more sensitive, there were some additional false positives, 
though at this problem is difficult to quantify. Regions with M.I. above 
$0.15$ consistently fit the FMRI data well, and simulations showed that
the particle filter performed extremely well in the face of significant
noise. In sum, the estimated BOLD output remained consistent despite
large swings in parameters.

% Pros/Cons
%% Cons
%%% Computation longer than SPM
%%% Harder to interpret
%% Pro
%%% Non-parametric
%%% Real time
%%% Full Posterior
%%% More intuitive
\section{Particle Filter Review}
The Particle Filter algorithm was originally designed for on-line parameter 
estimation. For this reason, there is no guarantee of optimality or even 
convergence for finite measurements. However, for the BOLD nonlinear ODE 
this is less of a concern than it might first appear to be. For this
particular problem there can be no guarantee of a global minimum, and although
other techniques guarantee a local minimum, the particle filter in some
sense can settle into multiple local minima.

@@

A more practical benefit with the particle filter is that it is mathematically
simple. An understanding of Bayesian statistics is all that is necessary to understand
how the particle filter works. This is in contrast to the Volterra based approach
of, which can be quite complex. Additionally very few assumptions
are needed for the particle filter; which reduces the risk of those assumptions
being violated. Fewer and more realistic assumptions make the particle 
filter more robust to unforeseen difficulties in FMRI data.

\chapter{Future Work}
\label{sec:FutureWork}
% Enhancements/Dehancements 
%% prior
%%% flatten prior
%%% Larger variance
%% Detrending
%%% \% difference using spline?
%%% "smart knots"
%%% DC gain as a parameter
%%% linearizing
%% Experimental Changes
%%% Variations in stimuli
%%% longer timeseries or artificial
%%% Automatic estimation of measurement error.
\section{Improving the Particle Filter}
\label{sec:Particle Filter Variations}

There are a few areas where future research may improve upon the current
work. The first is modifying the Prior 
Distribution. The choice in this paper to use the distribution calculated in
\cite{Friston2000}, may not be optimal and so is open to modification. The
second major area is in preprocessing, in particular the method used to remove
drift in the FMRI signal. This is an area of ongoing research and so future
algorithms could be strengthened by applying the latest technique rather than a 
spline. Third, the experiment itself may be modified to improve results.
And the last method is improving the , by using a more
advanced version.

Collapse parameters

\subsection{Prior Distribution}
The prior distribution used in this work is listed in \autoref{tab:Prior}, 
and is based on the findings of \cite{Friston2000}. Unfortunately, that
result depended on a quadratic approximation (Volterra Series) which has
not been extensively tested (at least not in a published work). As such,
the prior is probably not complete. Additionally, the current prior is
based only around the BOLD output, which is why it is capable of generating
good estimates of the BOLD signal. However, it would be advantagous to have
in-vivo estimates of the actual parameters. Better estimates could be found
using population studies with additional measurements as discussed in 
\autoref{sec:Sideways Measurements}. Changes to this prior could
boost the power of the BOLD particle filter algorithm described here.

Starting with a flattened
prior (by setting weights to be non-uniform after the initial particles have
been drawn) was not used in the final analysis. The reason for not flattening
the prior was a practical issue with weighting points in a 
7 dimensional joint distribution. Often differences in particle weights approached machine 
precision, meaning that the flattened prior was actually far from flat. Instead
the distribution became unpredictable, which made the results unpredictable. This
is quantization issue that will exist unless the prior were made to be deterministic
or the initial number particles was raised to computationally prohibitive levels.

Another improvement that could be made to the prior is increasing the initial variance.
This would allow the particle filter to learn a wider range of parameters; at the cost
of decreased convergence rate and support density. In tests, both real and simulated,
increasing the variance tended to allow the particle filter to over-smooth 
(ex \autoref{fig:badfit_param1}). In 
essence the particle filter tended to converge to a mean where the time constants
dominated the results by smoothing out all the peaks and troughs. The result is that
no other parameters alter the estimated BOLD signal. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/badfit_param1}
\caption{Large variance in time constants over-smoothing.}
\label{fig:badfit_param1}
\end{figure}
In \autoref{fig:badfit_param1}, the particle filter still had not fully converged, and
given more measurements it could still converge to a better estimate. Thus, increases
in the prior variance necessarily must be accompanied by more measurements,
although re-presenting data could be used to artificially extend the measurements.

\subsection{Detrending}
A large number of de-trending methods were tested, but all had some major problem. 
One viable method that could be utilized given the right experimental design is detrending
based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
constant to be added to the pre-processed signal for the BOLD model to fit properly.
The disadvantage of this approach is that it could hide long fall times by normalizing them
out. It also requires periodic breaks in the stimulus, which could reduce value of those
samples for fitting purposes.

Another potential way to deal with detrending is to 
linearize. This would use the delta between measurements for fitting rather than the 
direct value. This has the advantage of not requiring detrending and thus 
gives nearly raw data to the particle filter
for processing. The effectiveness of this method depends directly on the type of stimulus.
In an event driven stimulus with sparse stimuli this could be extremely effective because
the DC level has minimal data; however the longer high levels are held the less effective
this will be. On the other hand, splines will also reduce the plateaus, so its possible
linearizing may be just as effective. I found in tests that large drift-low white noise
type signals performed much better with a linearization approach, as one might expect. 
The results were equal to or worse in real data; but that was only for one particular
stimulus sequence.

A subtle difference that could be made in detrending is how the percent difference
is calculated. In this work the spline was subtracted, and then the result
was divided by the average of the initial signal. A more correct method may be dividing
by the spline value at that particular point. The reason for not dividing by the spline
at that point is that it could be less stable, and in a sense the trend has already
been removed. The difference is not particularly large though 
(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.

Finally, rather than adding a constant to the detrended data before
applying the particle filter, a DC gain parameter could be added to the model. In
tests, this could work well, however, it often did not. The problem with this
could simply be the addition of another degree of freedom without any increase in
the number of particles. Increasing the number of particles further though
can be computationally intractable. Thus, while this is in a sense the correct
solution, it is an impractical one.

\subsection{Experimental Changes}
\label{sec:Sideways Measurements}
One definite way of improving the results of the particle filter is more 
measurements. While increasing the sample rate of FMRI scanners may
not be possible, simultaneous measurements of volume and flow is
possible \cite{Hu2009}, albeit at 9T in a cat. Just one of those measurements
though would be extremely powerful when incorporated into the BOLD
model. By adding another measurement, especially for one of the other state
variables (like blood volume and blood flow), the
variance in the parameter estimates would certainly drop. Of course, more measurements
may be gained by simply performing longer FMRI tests, or concatenating several 
together. Even more basic, its possible to artificially increase the number of 
measurements by presenting them the time-series several times. 
This activity is similar to the process used in neural networks,
and gives the particle filter longer to converge to the optimal estimate. On the 
downside, this increases runtime and artificially reduces variance. 
Thus it does not necessarily improve the results. In tests, it did reduce the
final covariance but did not lead to better estimates.

Additionally, given the non-linearities in the system, differences in stimuli
can make a huge difference in the observability of parameters. The mentality 
for using a physiologically based nonlinear model for BOLD signal is to model 
those nonlinearities. Its logical then that nonlinear parameters may not
be identifiable when the input is primarily an impulse response. Therefore,
a wide range of inputs may shed further light on the parameters than found 
in this work.

As discussed in \autoref{sec:Methods Weighting Function}, choosing a weighting
function is difficult. Automatically estimating measurement error could improve
the quality of the particle filter results. Although I made some attempts to do
this, finding a generic, consistent solution is complex, and often depends on the
experimental design. This problem strikes at the very of heart of the difficulty
in analyzing FMRI, and so it is not likely to be solved soon.

% Future works
%% Using s as the input to other regions. 
%% Comparison of posterior of parameters
\section{Applications}
There are a number of advantages to the particle filter approach presented
here. In the past FMRI data has been analyzed strictly for determining
correlation between a stimulus and response. With this new method
the correlation is simply a means to determining the joint posterior distribution
of the parameters. While only regions that correlate with the input will 
be calculable, this method exploits that correlation to constrain the 
prior distribution of the parameters. The resulting distribution, while difficult 
to visualize because of the high-dimensionality, could nevertheless be 
correlated with neural pathologies. Despite the fact that the parameters
are under-determined, the final distribution is still a reduction in the
uncertainty of the parameters. This reduction in uncertainty naturally
contains information which may be exploited based on non-parametric 
statistical analyses.

Because of the limitations present in every image modality in neuroimaging,
its becoming increasingly clear that combining data from multiple sources
will be necessary to push neurology forward. In order to do so however, the
output of each source needs to fully represent what information that source
can provide. Combining the sources using Bayesian statistics is promising
yet often difficult because full probability distributions are hard to come by.
However, in this case, the particle filter provides a full posterior which
is extremely versatile. Therefore, future works will easily be able to
plug in data from multiple sources if they all output Bayesian posteriors.
For instance, if two different modalities have calculated the probability
distribution of neural efficiency, those two beliefs may be combined into
one conditional belief for the probability of neural efficiency.

An advantageous aspect of using a physiological model such as this, is that
it permits estimates of otherwise hidden parameters. In particular the BOLD
model gives an estimate of the value of the flow inducing signal, $s$. Having
this value available opens up new avenues for determining inter-regional
dependencies. All the regions for which the value of $s$ are trustworthy
correlate closely with the original stimuli, so determining the connection
between their values of $s$ would be less than edifying. Instead, the values
of $s$, which in some sense is the activation in a particular region, could
be used to drive other inputs. Thus, the particle filter could be re-run
with the time-series of a particular voxel's $s$ value as an additional stimulus.
In this way, it could be possible to determine chains of events. This is just
one possible benefit being able to determine the time course of the hidden
state variables present in the BOLD equations; the potential benefits of being
able to determine this information is limitless. Of course, improving the quality
of the parameter estimates is necessary if such information is to be fully
trusted.

>>>>

\chapter{Discussion}
\label{sec:Discussion}
\section{Review of Results}
% Overview of results
%% Parameters under-constrained
%% Output estimates good
There are two possible causes for the large variance in
parameter estimates. First, it could be that the particle filter is
incapable of learning the model. However, given the high quality
of BOLD estimates, this is almost certainly not the case. The other 
possible explanation is that the given stimulus is not sufficiently
varied to bring out all the properties of the system. While it is 
possible that this is the case, it is unlikely that even the best 
stimulus sequence could differentiate all the parameters. The fact
that \autoref{sec:VeryLongSim}, which used randomly spaced impulses,
still had a high correlation definitely indicates that the parameters
are ill-defined.

Thus there is little doubt that the parameters of the BOLD model are under
constrained. While unsurprising given the sensitivity analysis by Deneux et al.,
it is a notable conclusion \cite{Deneux2006}. Other methods
that depend on a point estimate of the parameters, such as least squares 
or Kalman filters (which uses the first two moments) are limited in their
capacity to estimate such parameters. While estimates of
the  BOLD signal may still be correct, the 
underlying parameters and state variables cannot be described without using
a joint probability distribution function. In this sense, particle 
filters represent an important step forward in BOLD parameter 
estimation. Representing uncertainty with a mean
and variance is insufficient; so using a particle filter
or Bayesian estimate of the posterior is not simply an enhancement, 
but a necessary precaution.

Because of the ill-defined nature of the parameters, the results 
of Friston et al. may have been wider than true parameter distribution
\cite{Friston2002}. Future works that attempt to calculate parameters
of the BOLD model should be run multiple times to ensure consistency,
which most likely cannot be attained. This result is particularly
troublesome given the number of papers that use those parameters for
a prior distribution.

As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information, closely resembled the results of SPM. While the activation
tests were more sensitive, there were some additional false positives, 
though the problem is difficult to quantify. Regions with M.I. above 
$0.15$ consistently fit the FMRI data well, and simulations showed that
the particle filter performed extremely well in the face of significant
noise. In sum, the estimated BOLD output remained consistent despite
large swings in parameters.

% Pros/Cons
%% Cons
%%% Computation longer than SPM
%%% Harder to interpret
%% Pro
%%% Non-parametric
%%% Real time
%%% Full Posterior
%%% More intuitive
\section{Particle Filter Review}
The Particle Filter algorithm was originally designed for on-line parameter 
estimation. For this reason, there is no guarantee of optimality or even 
convergence for finite measurements. However, for the BOLD nonlinear ODE 
this is less of a concern than it might first appear to be. For this
particular problem there can be no guarantee of a global minimum, and although
other techniques guarantee a local minimum, tests show that the particle
filter did converge relatively quickly \autoref{sec:VeryLongSim}.

@@

A more practical benefit with the particle filter is that it is mathematically
simple. An understanding of Bayesian statistics is all that is necessary to understand
how the particle filter works.  Additionally very few assumptions
are needed for the particle filter. Few assumptions  reduces the risk of those assumptions
being violated. Fewer and more realistic assumptions also make the particle 
filter more robust to unforeseen difficulties in FMRI data.

%\section{Particle Filter Options}
%There are a large number of options available in the particle filter, that 
%were of little use. 
%\section{Didn't Work}
%A more subtle difference that could be made in detrending is how the percent difference
%is calculated. In this work the spline was subtracted, and then the result
%was divided by the average of the initial signal. A more correct method may be dividing
%by the spline value at that particular point. The reason for not dividing by the spline
%at that point is that it could be less stable, and in a sense the trend has already
%been removed. The difference is not particularly large though 
%(dividing by 1020 rather than 1000 for instance) as long as the spline didn't have heavy swings.

%Finally, rather than adding a constant to the detrended data before
%applying the particle filter, a DC gain parameter could be added to the model. In
%tests, this could work well, however, it often did not. The problem with this
%could simply be the addition of another degree of freedom without any increase in
%the number of particles. Increasing the number of particles further though
%can be computationally intractable. Thus, while this is in a sense the correct
%solution, it is an impractical one.

%Starting with a flattened
%prior (by setting weights to be non-uniform after the initial particles have
%been drawn) was not used in the final analysis. The reason for not flattening
%the prior was a practical issue with weighting points in a 
%7 dimensional joint distribution. Often differences in particle weights approached machine 
%precision, meaning that the flattened prior was actually far from flat. Instead
%the distribution and results became unpredictable. This
%is quantization issue that will exist unless the prior were made to be deterministic
%or the initial number particles was raised to computationally prohibitive levels.

%Another improvement that could be made to the prior is increasing the initial variance.
%This would allow the particle filter to learn a wider range of parameters; at the cost
%of decreased convergence rate and support density. In tests, both real and simulated,
%increasing the variance tended to allow the particle filter to over-smooth 
%(ex \autoref{fig:badfit_param1}). In 
%essence the particle filter tended to converge to a mean where the time constants
%dominated the results by smoothing out all the peaks and troughs. The result is that
%no other parameters alter the estimated BOLD signal. 
%\begin{figure}
%\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=7cm]{images/badfit_param1}
%\caption{Large variance in time constants over-smoothing.}
%\label{fig:badfit_param1}
%\end{figure}
%In \autoref{fig:badfit_param1}, the particle filter still had not fully converged, and
%given more measurements it could still converge to a better estimate. Thus, increases
%in the prior variance necessarily must be accompanied by more measurements,
%although re-presenting data could be used to artificially extend the measurements.

%Even more basic, its possible to artificially increase the number of 
%measurements by presenting them the time-series several times. 
%This activity is similar to the process used in neural networks,
%and gives the particle filter longer to converge to the optimal estimate. On the 
%downside, this increases runtime and artificially reduces variance. 
%Thus it does not necessarily improve the results. In tests, it did reduce the
%final covariance but did not lead to better estimates.

\chapter{Future Work}
\label{sec:FutureWork}
% Enhancements/Dehancements 
%% prior
%%% flatten prior
%%% Larger variance
%% Detrending
%%% \% difference using spline?
%%% "smart knots"
%%% DC gain as a parameter
%%% linearizing
%% Experimental Changes
%%% Variations in stimuli
%%% longer timeseries or artificial
%%% Automatic estimation of measurement error.
\section{Algorithm Improvements}
\label{sec:Particle Filter Variations}
There are a few areas where future research may improve upon the current
work. The first is modifying the prior 
distribution. The prior distribution used in this work is listed in \autoref{tab:Prior}, 
and is based on the findings of Friston et al. \cite{Friston2000}. Unfortunately, that
result depended on a quadratic approximation (Volterra Series) which has
not been extensively tested (at least not in a published work). 
Additionally, the current prior is
based only around the BOLD output, which is why it is good at generating
estimates of the BOLD signal, but weak in estimating parameters. 
The best solution would be to have
in-vivo estimates of the actual parameters, although this
is unlikely to happen. Instead, better estimates could be found
using population studies with additional measurements as discussed in 
\autoref{sec:Sideways Measurements}. Better knowledge of the prior distribution
would make it possible to decrease variance of certain 
parameters, reducing the mobility of the parameters.

Another major area for improvement is removing drift. 
A large number of detrending methods were tested, but every method was in some way limited. 
One viable method that could be utilized, given the right experimental design, is detrending
based on areas of low activity. This has the advantage that it wouldn't require an arbitrary
constant to be added to the pre-processed signal for the BOLD model to fit properly.
The disadvantage of this approach is that it could hide long fall times by normalizing them
out. It also requires periodic breaks in the stimulus, which could reduce value of those
samples for fitting purposes.

In terms of dealing with drift, linearizing is another possibility.
This would use the delta between measurements for fitting rather than the 
direct value. This has the advantage of not requiring detrending and thus 
gives nearly raw data to the particle filter
for processing. The effectiveness of this method depends directly on the type of stimulus.
In a test with rapid impulse stimuli, this could be extremely effective because
the DC level has minimal data; conversely prolonged flat level levels make this less effective.
In tests I found that large drift-low white noise
type signals performed much better with a linearization approach, as one might expect. 
For the stimulus sequence used in \autoref{sec:RealData}, the results tended to be worse
than using spline detrending.

As discussed in \autoref{sec:Methods Weighting Function}, choosing a weighting
function is difficult, and the optimal solution varies based on the input.
Automatically estimating measurement error could improve
the quality of the particle filter results. Although I made some attempts to do
this, finding a generic, consistent solution is complex, and often depends on the
experimental design. Part of the problem is that SNR varies greatly across
the brain, and there is no way to actively measure noise without also knowing
the underlying signal. One possible solution is having the particle filter automatically
set the weight based on the total particle weight, or on the prevalence of resampling.

\subsection{Experimental Changes}
\label{sec:Sideways Measurements}
One definite way of improving the results of the particle filter is additional
measurements. While increasing the sample rate of FMRI scanners may
not be possible, simultaneous measurements of volume and flow is
possible, albeit at 9T in a cat \cite{Hu2009}. Just one of those measurements
though would be extremely powerful when incorporated into the BOLD
model. By adding another measurement, the
variance in the parameter estimates would significantly drop. The fact
that such a measurement would be closer to the true stimulus would make it
all the more powerful. 
A more conventional method of adding measurements is to 
simply perform longer FMRI tests. Although this wouldn't be 
groundbreaking, it would certainly increase the ability of the 
particle filter to develop the posterior distribution. 
It may also be possible to concatenate multiple FMRI tests together.
It should be possible to link the multiple runs simply through
the final parameter distribution from the previous run.

Given the non-linearities in the system, differences in stimuli
could make a large difference in the observability of parameters. The mentality 
for using a physiologically based nonlinear model for BOLD signal is to model 
those nonlinearities. Its logical then that certain nonlinear parameters may not
be identifiable when the input is primarily an impulse response. It 
has been reported that short responses are disproportionately large 
in FMRI data \cite{Miller2001, Deneux2006}. Therefore,
a wide range of activation hold time may shed further light on 
the parameters' distribution as well as the validity of the BOLD model.

% Future works
%% Using s as the input to other regions. 
%% Comparison of posterior of parameters
\section{Future Applications}
There are a number of advantages to the particle filter approach presented
here. In the past, FMRI data has been analyzed strictly for determining
correlation between a stimulus and response. With this new method
the correlation is simply a means to determining the joint posterior distribution
of the parameters. While only regions that correlate with the input will 
be calculable, this method exploits that correlation to constrain the 
prior distribution of the parameters. The resulting distribution, while difficult 
to visualize because of the high-dimensionality, could nevertheless be 
correlated with neural pathologies. In spite of the fact that the parameters
are under-determined, the final distribution is still a reduction in the
uncertainty of the parameters. The availability of a full posterior distribution
opens up many avenues for further inquiry, for instance
to compare normal vs. symptomatic
populations. This would be especially useful in patients whose symptoms
do not include structural changes in brain. This would effectively be
a way to differentiate the way the brain operates.

Because of limitations present in every imaging modality,
its becoming increasingly clear that combining data from multiple sources
will be necessary to push Neurology forward. In order to do so however, the
output of each source needs to fully represent what information that source
can provide. Combining the sources using Bayesian statistics is promising
yet often difficult because full probability distributions are hard to come by.
However, in this case, the particle filter provides a full posterior which
is extremely versatile. Therefore, future works will easily be able to
plug in data from multiple sources if they all output Bayesian posteriors.
For instance, if two different modalities have calculated the probability
distribution of neural efficiency, those two beliefs may be combined into
one conditional belief for the probability of neural efficiency.

An advantageous aspect of using a physiological model such as this, is that
it permits estimates of otherwise hidden parameters. In particular the BOLD
model gives an estimate of the value of the flow inducing signal, $s$. Having
this value available opens up new avenues for determining inter-regional
dependencies. The values of $s$ could serve as a proxy for activation in 
that region and thus could be used to drive other inputs regions' $u(t)$. 
Thus, the particle filter could be re-run
with the time-series of a particular voxel's $s$ value as an additional stimulus.
In this way, it could be possible to determine chains of events. This is just
one possible benefit being able to determine the time course of the hidden
state variables present in the BOLD equations; the potential benefits of being
able to determine this information is limitless. Of course the quality of these
estimates will continue to improve with the model priors.
